|

### 2021-08-15 dissertation/msc_dissertation.Rmd
Cleaning up the code a bit

---
title: |
  | __How analysis strategy affects analysis results__
subtitle: |
  | *Assessing the median effect robustness in @silberzahn2018 through model specification*
  |
  | Master's thesis - DRAFT
author: "Sebastian Ploner, sploner1@sheffield.ac.uk, University of Sheffield"
linestretch: 1.5
output:
    pdf_document:
        includes:
            in_header: header.tex
        extra_dependencies: ["flafter"]
bibliography: references.bib
csl: apa.csl
header-includes:
  - \usepackage[ruled,vlined]{algorithm2e}
---

@@

# Download and install missing packages
requiredPackages = c("here", "data.table", "tidyverse", "PupillometryR", "cowplot", "knitr", "kableExtra", "papeR", "DiagrammeR")

@@

# Load data
dat = data.frame(fread(here::here("data", "3_prepared_data.csv")))

@@

 Taken together, researchers make many analytical decisions throughout conduction a study which are know as *researcher degrees of freedom*. Crowdsourcing analysis presents a strategy to understand their induced variation, but is impractical due to its substantial time and personnel requirements. Multiverse analysis present a promising, low-personnel, but computationally intense alternative to assessing the scope this variation. Despite observing substantial variation in crowdsourced data analysis projects little is known about the extend to which these approaches cover the full range of possible results. In this project, using methods of multiverse analysis I will define the results space of @silberzahn2018 to estimate study's representativeness. Based on these insights I aim at contributing to further understanding the value, limitations and mechanisms of the multiverse approach.

# Analysis
## Analysis plan
     The project's objective is to define the results space of the @silberzahn2018 to the estimate study's representativeness. The results space refers to the numerical interval between the lowest and highest possible outcome. It is created by running every possible analytical strategy. Hence, it is closely related to assessing an effect's robustness. @patel2015 developed a standardised approach to assessing an effect's robustness which they termed assessing the *vibration of effect*. I will therefore use this standardised approach to define the results space. The approach essentially comprises two parameters: the statistical models and the covariates (or control variables). In @silberzahn2018 the analysts used numerous different statistical models like multiple linear regression, mixed-model logistic regression or Bayesian logistic regression. In total there were 29 different modelling approaches. Additionally, each team used a different set of covariates. Across all teams there 15 covariates used. This gives $2^n$ i.e., $2^{15} = 32,768$ possible combinations of covariates. Adding all modelling possibilities to the equation gives a total of $29 * 2^{15} = 950,272$ combinations to run. This number exceeds the project's scope, hence, it needs to be reduce to a more manageable count. I am, therefore, focusing on one modelling approach. In particular, on the approach that produced the median outcome of all analyses. The median, being the middle number of any set of values, is a reasonable starting point in order to estimate the results space. Nevertheless, even focusing on one analytical approach still leaves $1 * 2^{15} = 32,768$ possibilities. Due to computational limitations I will first run a sample of 200, followed by a sample of 1,000 combinations. If there's a substantial difference between the median and the spread I will run more models, if not, I will make the assumption that 1,000 reasonably estimates all ~33k combinations.

 In @silberzahn2018 the median outcome was produced by @team23 which will hereinafter be referred to as *team 23* due to it's designated team number in the original study. *Team 23* first transformed the data and then conducted a mixed-model logistic regression. The first step of this project will be replicating the transformation and the analysis. Replicating other researchers analyses has the benefit of checking their work and, if results are indeed replicated, increasing confidence in them. It also ensures that this project has the same starting point as *team 23* did. Given the team made all their scripts publicly available I expect this step to be straightforward. Next is drawing a random sample of covariates, without replacement. "Without replacement" ensures all covariate combinations are unique in the sample. The covariates will be appended to the base (or core) variables. Base variables are those variables that are primarily assessed to answer the research question. In this case whether a football player's skin colour effects the odds of being sent off the field. *Team 23* defined two interaction terms as the base: "skin tone *X* implicit bias" and "skin tone *X* explicit bias". (The variables are described in the "data" section.) Hence, all models will have the following structure:
$$
\begin{aligned}
RedCard = SkinTone*ImplicitBias + SkinTone*ExplicitBias + CovariateCombination_{i}
\end{aligned}
$$ 
While running each model, the relevant statistics will be extracted. Those are the coefficient (or effect) of skin tone, its standard error, test-statistic and p-value. Just as *team 23* did I will then calculate the 95% confidence intervals (CI). The estimates and their CIs are then transformed to odds ratios (OR) through exponentiating them to the power of two. Odds ratios quantify the strength of association between two variables. If greater than one the dependent variable is more likely to occur given the independent variable, if lower than one it less likely. The odds ratios will be visualised to assessed through a conventional specification curve plot, a rain-cloud plot and scatter plot. The following describes them in more detail:

(i) The specification curve plot was developed to be a descriptive plot i.e., raw data is being plotted without any aggregation done. Its objective is describe the results space while allowing the reader to identify the specified covariate for each outcome [@simonsohn2020]. It therefore has two vertically stacked components. The top part shows a sorted scatter plot. On the horizontal axis are the models specifications and on the vertical axis is the outcome measure. The points are sorted from lowest to highest outcome measure. This way the lowest outcome measure is on the bottom left corner and the highest in the top right one. The bottom part of the plot represents are table. Just like the top the horizontal axis hold the model specifications, the vertical axis lists the covariates. This arrangement allows each cell to specify the presence or absence for each covariate in a given specified model. For the top and bottom plot to work together it is imperative that the horizontal axes are identically arranged. If so, for every point in the top plot (i.e., for every outcome) the specified covariates can be seen in the bottom plot. For an example check the results section. The issue, however, is that if too many specifications are on the horizontal axis it becomes very hard to identify specific covariates. Hence, I also included the rain-cloud and scatter plot.

(ii) The rain-cloud plot was developed by @allen2021. Its objective is to give an unbiased, transparent view on the raw data. It therefore combines a density, box and scatter plot. If stacked vertically from top to bottom, respectively, they look like a cloud with rain drops, hence, its name. The advantage of this plot is to be able to assess the raw data including key statistics (median, 25th and 75th quartiles and CIs) as well as the probability density all in one succinct plot. Given that this project seeks to define and describe a results space the rain-cloud plot the ideal graph to do this. However, the plot neither allows to identify the covariates giving rise to the outcome nor their specified effects. (For an example check the results section.) Therefore a third plot is developed which is to be assessed in tandem with this plot.

(iii) The third and final plot is a sorted scatter plot. While there is nothing special about the plot itself its content is important. Given the very high number of models that are run, it does not make sense to assess each model and its effects in isolation. Instead the effects for each covariates neeed to be aggregated. Hence, the goal of this plot is to viusalise the specified effect for each covariate. Prior to doing so, the coviates first need to be aggregated and their specified effects need to be calucualted. To this end, each covariate is recoded into a binary factor: included in the model yes/no. This newly defined factors are than used as an independent variables in an ANOVA. The previously calcualted odds rations are used as dependent variables. The estimates of the fitted model are the specified effect for each covariates. These are visualised similarly to the top part of the specification curve plot. The covariates are on the horizental axis, will the odds ratios are on the vertical axis. Here too, the lowest oucome is in the bottom left corner and the highest outcome in the top right one. Additionally the point are coloured indicating a statistically significant effect (or non-signifcant).

## Code
     The code is based on @haessler2020 and @patel2015. The former provided the operational backbone of the code. The latter the idea for the algorithm. R's random number generator is first fixed to be able to reproduce the outcomes. Then, all the packages are detached and only the relevant ones are loaded to prevent other functions from interfering. If the required packages are not installed, the code will do so. Next, the prepared data is loaded using the "Here" package [@here] for operating systems independence and the "data.table" package for faster importing of the data [@datatable]. (The data has been prepared by running the "disaggregate_v3.py" script of *team 23*; for more detail see the "data" section.) After having loaded the data the variables are defined as dependent, base (or core) variables and covariates. This includes defining the variables' types: numeric, factor or random effect. Next, a matrix is created established all possible combination of covariates (including base variables). Each row represents one combination, the columns represent the variables. The cells are set to TRUE or FALSE to indicate a variables presence or absence, respectively. Thus, the matrix structure looks as the following:

$$\begin{bmatrix}
\text{dv} & \text{base var}_{1} & \text{base var}_{2} & \text{base var}_{3} & \text{covar}_{1} & \text{covar}_{2} & \text{covar}_{...} & \text{covar}_{15} \\
\text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{FALSE} & \text{FALSE} & \text{...} & \text{FALSE} \\
\text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{FALSE} & \text{...} & \text{FALSE} \\
\text{...} & \text{...} & \text{...} & \text{...} & \text{...} & \text{...} & \text{...} & \text{...} \\
\text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{...} & \text{TRUE} \\
\end{bmatrix}$$ \newline

 Using the *apply* function, looping over each row, the variables are pasted together thereby creating the corresponding formulas which are then saved as a vector and are appended as column to the aforementioned matrix.

 *Team 23* ran a mixed-model logistic regression analysis using the "lme4" package more specifically, its "glmer" function [@lme4]. The function only works if a random structure is included in the formula. As all possible covariate combinations are explored, a random structure is not included in every formula. Therefore, before running the models, the formulas need to separated based on including a random structure or not. The random structures were defined at the beginning of the code using the following syntax: "(1|*covariate*)". Hence, this pattern is also used to identify random structures through the "grepl" function. For the non-random structure formulas the regular "glm" function is run. Both models, mixed-effect and "regular", are estimated using Maximum Likelihood estimation to ensure the outcomes are as comparable as possible. *Team 23* had also set the "nAGQ" parameter to zero (for the mixed-model) this sacrificed parameter estimation accuracy for speed. The same was done here to stay as close to the original analysis as possible. Finally, 1,000 formula samples were drawn. 75% of all formulas include at least one random structure this proportion is reflected in the sample. Furthermore, the samples are drawn without replacement i.e., every formula is unique.

 Running the models yielded another challenge. Not only were the running times per model at times above 30 minutes, storing the model summaries also required significant memory. For instance, testing the code with a minimal sample of 10 (random structure) formulas yielded an analysis output larger than 400MB. This does not scale well, hence, a wrapper function was build. It's purpose was to extract the estimates, standard errors, test-statistics and p-values for the variable of interest (in this case: skin tone). This effect was substantial. The final outcome object for 1,000 models was reduced to 0.3MB(!).

 The models were run using the "lapply" function more specifically, the paralellised version thereof including progress bars ("pbmcapply", @pbmcapply). Parallisation was necessary in order to reduce running time. The standard R is set to use one core, only. Parallelising the processes to six cores decreased the running time by six-fold. The results of each model iteration was appended to a list. Hence, the outcomes were two lists, one from the mixed-models, and one from the "regular" models. These lists were transformed to data frames and merged. Moreover, the p-values was coded into significant and non-significant ($\alpha = 0.05$). The odds ratios as well as their confidence intervals were calculated (the same way *team 23* did it). The resulting data frame was save as a comma-separated values (CSV) file.

@@

... \newline
*Explaining why not adjusting for multiple comparison--in progress* \newline
... \newline

To better understand the specified effects of the covariates I ran an ANOVA where each covariate was coded into included yes/no. The above retrieved effects where used outcome measure. The results were stored and saved as a separate data frame. As a final step the graphs were created using the "tidyverse" (specifically, "ggplot", @tidyverse), "PupillometryR" [@PupillometryR] and "cowplot" [@cowplot] packages. \newline

... \newline
*Explaining how the graphs were created including necessary data transformation needed--in progress.* \newline
... \newline

The code is written in R (Version 1.4.1103) on macOS Big Sur (Version 11.4) and can be retrieved from my [*GitHub repository*](https://github.com/sebastianplnr/msc_dissertation_project).

## Data origin and preparation
Outline:

* The same dataset as in @silberzahn2018 is used.

* What is the data about

* Data origin i.e., how did silberzahn et al. (2018) come up with the data set

* how was the data processed by team 23

* summary statistics of the data set

```{r, echo = FALSE}

# Setting variable class
dat$player = as.factor(dat$player)
dat$specific_pos = as.factor(dat$specific_pos)
dat$club = as.factor(dat$club)
dat$league_country = as.factor(dat$league_country)
dat$ref = as.factor(dat$ref)
dat$ref_country = as.factor(dat$ref_country)

# Building a summary table
kable(papeR::summarise(dat,
                       type = "numeric",
                       variables = c("height_cm", "weight_kg", "age_yrs", "games", "goals", "victories", "ties", "all_reds", "player_cards_received",
                                     "ref_cards_assigned", "skin_tone_num", "imp_bias", "exp_bias")),
      format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = "scale_down")

@@

\newpage
## Outcome I: vibration of effect
## Outcome II: specified covariate effects
## Outcome III: specification curve

\newpage
# Discussion

>>>>

---
title: |
  | __How analysis strategy affects analysis results__
subtitle: |
  | *Defining the results space in @silberzahn2018 through model specification*
  |
  | Master's thesis - DRAFT
author: "Sebastian Ploner, sploner1@sheffield.ac.uk, University of Sheffield"
linestretch: 1.5
output:
    pdf_document:
        includes:
            in_header: header.tex
        fig_caption: yes
urlcolor: blue
bibliography: references.bib
csl: apa.csl
header-includes:
  - \usepackage[ruled,vlined]{algorithm2e}
---

@@

# Download and install missing packages
requiredPackages = c("here", "data.table", "tidyverse", "PupillometryR", "cowplot", "knitr", "DiagrammeR")

@@

# Load data
## prepared data
dat = data.frame(fread(here::here("data", "3_prepared_data.csv")))

## Outcomes of running the 1,000 models
dat_models = data.frame(fread(here::here("data", "4_model_outcome_data.csv")))

# 
dat_models_specifications = data.frame(fread(here::here("data", "4_model_outcome_data_specifications.csv")))

## Outcome of running the "meta-ANOVA" to get the covariates' specified effects
dat_covariates = data.frame(fread(here::here("data", "4_covariate_effect_data.csv")))

# Adjusting class for proper visualisation
dat_models$below_alpha = factor(dat_models$below_alpha)

dat_covariates$impact_names = factor(dat_covariates$impact_names)
dat_covariates$below_alpha = factor(dat_covariates$below_alpha)

@@

 Taken together, researchers make many analytical decisions throughout conduction a study which are know as *researcher degrees of freedom*. Crowdsourcing analysis presents a strategy to understand their induced variation, but is impractical due to its substantial time and personnel requirements. Multiverse analysis present a promising, low-personnel, but computationally intense alternative to assessing the scope this variation. Despite observing substantial variation in crowdsourced data analysis projects little is known about the extend to which these approaches cover the full range of possible results. In this project, using principles of multiverse analysis I will define the results space of @silberzahn2018. By doing so I aim at further understanding how a specific analysis strategy affects its result. To my knowldge this is the first project runing not only all *plausible* but all *possible* analytical strategies to a more granular view of the results space.

# Analysis
## Analysis plan
     The project's objective is to define the results space of the @silberzahn2018 to draw conclusions about the mechanisms of multiverse analysis. The results space refers to the numerical interval between the lowest and highest possible outcome. It is created by running every possible analytical strategy. Hence, it is closely related to assessing an effect's robustness. Robustness refers to an effect's consistency under different model specifications. @patel2015 developed a standardised approach to assessing an effect's robustness which they termed assessing the *vibration of effect*. I will therefore use this standardised approach to define the results space. The approach essentially comprises two parameters: the statistical models and the covariates (or control variables). In @silberzahn2018 the analysts used numerous different statistical models like multiple linear regression, mixed-model logistic regression or Bayesian logistic regression. In total there were 29 different modelling approaches. Additionally, each team used a different set of covariates. Across all teams there 15 covariates used. This gives $2^n$ i.e., $2^{15} = 32,768$ possible combinations of covariates. Adding all modelling possibilities to the equation gives a total of $29 * 2^{15} = 950,272$ combinations to run. This number exceeds the project's scope, hence, it needs to be reduce to a more manageable count. I am, therefore, focusing on one modelling approach. In particular, on the approach that produced the median outcome of all analyses. The median, being the middle number of any set of values, is a reasonable starting point in order to estimate the results space. Nevertheless, even focusing on one analytical approach still leaves $1 * 2^{15} = 32,768$ possibilities. Due to computational limitations I will first run a sample of 200, followed by a sample of 1,000 combinations. If there's a substantial difference between the median and the spread I will run more models, if not, I will make the assumption that 1,000 reasonably estimates all ~33k combinations.

 In @silberzahn2018 the median outcome was produced by @team23 which will hereinafter be referred to as *team 23* due to it's designated team number in the original study. *Team 23* first transformed the data and then conducted a mixed-model logistic regression. The first step of this project will be replicating the transformation and the analysis. Replicating other researchers analyses has the benefit of checking their work and, if results are indeed replicated, increasing confidence in them. It also ensures that this project has the same starting point as *team 23* did. Given the team made all their scripts publicly available I expect this step to be straightforward. Next is drawing a random sample of covariates, without replacement. "Without replacement" ensures all covariate combinations are unique in the sample. The covariates will be appended to the base (or core) variables. Base variables are those variables that are primarily assessed to answer the research question. In this case whether a football player's skin colour affects the odds of being sent off the field. *Team 23* defined two interaction terms as the base: "skin tone *X* implicit bias" and "skin tone *X* explicit bias". (The variables are described in the "data" section.) Hence, all models will have the following structure:
$$
\begin{aligned}
RedCard = SkinTone*ImplicitBias + SkinTone*ExplicitBias + CovariateCombination_{i}
\end{aligned}
$$ 
While running each model, the relevant statistics will be extracted. Those are the coefficient (or effect) of skin tone, its standard error, test-statistic and p-value. Just as *team 23* did I will then calculate the 95% confidence intervals (CI). The estimates and their CIs are then transformed to odds ratios (OR) through exponentiating them to the power of two. Odds ratios quantify the strength of association between two variables. If greater than one the dependent variable is more likely to occur given the independent variable, if lower than one it less likely.

 These odds ratios will be visualised to assessed through a conventional specification curve plot, a rain-cloud plot and scatter plot. The following describes the three graphs in more detail:

(i) The specification curve plot was developed to be a descriptive plot i.e., raw data is being plotted without any aggregation done. Its objective is to describe the results space while allowing the reader to identify the specified covariates for each outcome [@simonsohn2020]. It therefore has two vertically stacked components. The top part shows a sorted scatter plot. Its horizontal axis shows the models specifications and its vertical axis displays the outcome measure. The scatter points are sorted from lowest to highest outcome measure. This way the lowest outcome measure is on the bottom left corner and the highest in the top right one. The bottom part of the plot represents a table. As in the top plot the horizontal axis holds the model specifications, the vertical axis lists the covariates. This arrangement allows each cell to specify the presence or absence for each covariate in a given specified model. For the top and bottom plot to work together it is imperative that the horizontal axes are identically arranged. If so, for every point in the top plot (i.e., for every outcome) the specified covariates can be seen in the bottom plot. For an example check Figure \ref{fig:sca_plot}. As the current project has 1,000+ specifications it becomes very hard to identify specific covariates given the space limitations. Hence, I to had come up with different more intuitive ways of visualisation.

(ii) The rain-cloud plot was developed by @allen2021. Its objective is to give an unbiased, transparent view on the raw data. It therefore combines a density, box and scatter plot. If stacked vertically from top to bottom, respectively, they look like a cloud with rain drops, hence, its name. The advantage of this plot is to be able to assess the raw data including key statistics (median, 25th and 75th quartiles and CIs) as well as the probability density all in one succinct plot. The current project seeks to define and describe a results space the rain-cloud plot is, therefore, the ideal choice of graphs. (For an example check the results section.) Unfortunately, the plot does not allow to identify the covariates giving rise to the outcomes or their specified effects. As the project also seeks to understand the mechanisms of multiverse analysis it important to get insights about the specified covariates effects. Therefore a third plot is developed which is to be assessed in tandem with this plot.

(iii) The third and final plot is a sorted scatter plot. While there is nothing special about the plot itself its content is important. Given the very high number of models that are run, it does not make sense to assess each model and its effects in isolation. Instead the effects for each covariates need to be aggregated. Hence, the goal of this plot is to visualise the specified effect for each covariate. Prior to doing so, the covariates first need to be aggregated and their specified effects need to be calculated. To this end, each covariate is recoded into a binary factor: included in the model yes/no. This newly defined factors are than used as an independent variables in an ANOVA. The previously calculated odds rations are used as dependent variables. The estimates of the fitted model are the specified effect for each covariates. These are visualised similarly to the top part of the specification curve plot. The covariates are on the horizontal axis, will the odds ratios are on the vertical axis. Here too, the lowest outcome is in the bottom left corner and the highest outcome in the top right one. Additionally the point are coloured indicating a statistically significant effect (or non-significant).

 Taken together, the current project's objective is to define the results space of @silberzahn2018 in order to make inferences about the underlying mechanisms of multiverse analysis. The results space is created by running all possible model and covariate combinations. Applied to @silberzahn2018 this means running roughly 1M models. Due to time and computational limitations, I am focusing on the model that produced the median outcome of all analytical strategies. Hence, I will run a mixed-model logistic regression with 1,000 randomly sampled covariate combinations. For each model relevant statistics will extracted and visualised. The conventional specification curve plot does not deal well with a large number of model specifications. Therefore, I will produce a rain-cloud and scatter plot, which visualise the results space (and its distribution) as well as the specified covariate effects, respectively. For the latter, I will first have to run an additional statistical model to calculate the specified effects. These graphs will give insight into the results space, its building blocks and ultimately into the mechanisms of multiverse analysis.

## Code
     This section provides a brief overview of the code and explains the reasoning behind it. As mentioned in the analysis section the first step was replicating the data transformation and analysis of *team 23* [@team23]. The team made their project folder publicly available. It contains three scripts relevant to this project: the data exploration, transformation and analysis. After duplicating their project folder on my local hard drive all scripts run without issues (only the working directories needed adjustment). The data exploration included the reasoning behind transforming the data and some cleaning. The data transformation restructured the data into a more intuitive format (more details on the topics of exploration and transformation are in the data section). The analysis script prepared the data by assigning variables classes (factors, numerical or boolean) and standardised a few variables i.e., they were centred around the mean. Finally, the models were specified. The team ran both frequentist and Bayesian models, however, this project covers the frequentist approach, only.

 My code is based on @haessler2020 and @patel2015. The rough structure of the code is shown in the pseudocode table (Figure X). Here, I am briefly motivating two analytical choices I made. For a closer look please check the commented code itself:

(i) Choosing the function to run the statistical model. *Team 23* ran a mixed-model logistic regression. This model has two relevant components to this explanation: fixed effects and random effects. Fixed effects are those that are consistently observed in different situations because the construct is of direct interest to the research question. In this case, it is, for example, skin tone. Independent of the player, the game or the league skin tone is always observed. Their counter parts are random effects which change between situations. A player, for example, is just one "unit" a measurement, and is himself not directly relevant to answering the research question. The function used by *team 23* requires to specify a random effect. However, given the project seeks to sample from all possible covariate combinations, a random effect is not always included. Hence, the "non-random" counter part to this function needs to be utilised. For this project, it was important to check that both functions use the same estimation method (maximum likelihood or restricted maximum likelihood estimation) to calculate the statistics. It was important to check to ensure the outcomes are comparable.

(ii) No multiple comparison. This is based on the fact that in statistics null hypotheses are one rejected if the probability of finding a false positive is below 5% (known as the significance level, $\alpha$). The more statistical tests a run, the more likely it is that one of those tests is indeed a false positive. Hence, the are techniques that account of this issue. It is usually good practice to account for those "multiple comparisons". Here, I refrained from doing so though. This project seeks to simulate multiple researchers running different models. Therefore, for the the purpose of this project all models are run independently by individual researchers. Those would not have knowledge of the other analyses and, hence, would not account for them. To main the highest possible ecological validity, I therefore did not account of multiple comparisons.

@@

 The code is written in R (Version 1.4.1103) on macOS Big Sur (Version 11.4) and can be retrieved from my [GitHub repository](https://github.com/sebastianplnr/msc_dissertation_project). The code from *team 23* [@team23] can be retrieved from their [OSF repository](https://osf.io/akqt4/). Following R packages were used *here 1.0.1* [@here], *data.table 1.14.0* [@datatable], *tidyverse 1.3.1* [@tidyverse], *lme4 1.1-27.1* [@lme4], *pbmcapply 1.5.0* [@pbmcapply], *PupillometryR 0.0.3* [@PupillometryR] and *cowplot 1.1.1* [@cowplot].

## Data description and preparation
     The data was retrieved from @silberzahn2018. It contains information about football players, their encounters with referees and the received cards (yellow, yellow/red and red). Moreover, a player's position, age, club, league country, victories, ties, defeats and goals. In addition, a skin tone rating based on two independent raters was included. The referees are numerically coded to protect their identify. The referees origin countries are also included as well as an implicit and explicit racism bias scores for their respective countries. The exploratory data analysis (EDA) of *team 23* showed that each row of the dataset represents a unique player-referee combination listing all their encounters as well as the couple's total number of received/assigned cards. The team states that it preferred a different data format. More specifically, a format where each player-referee encounter is reflected by one row. This way each encounter had a maximum of one red card. To achieve this format the data had to be "disaggregated" i.e., transformed. Further EDA showed that receiving a red card is highly unlikely, 0.008%, i.e., data is highly skewed. This property informed the teams decision to use a type of logistic regression (a statistical modelling technique equipped to deal with skewed/binary distributions). Finally, the team excluded all referee that did have at least 22 encounters. Their reasoning was a game includes (at least) 22 players, 11 players per team, if this were not the case referee most likely did not officiate in one for the four leagues of interest (England, Germany, Italy and Spain). This step excluded roughly 66% of the referees but retained 97.4% of the cases. The final dataset used for the current project contained 335,537 observations and 19 variables. Figure \ref{fig:var_plots} shows the properties of the variables relevant to current project's analysis \newline

```{r var_plots, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "A) Showing the strong skew of receiving no red to receiving a red card. B) Showing most players had a lower skin tone rating i.e., most players were on the whiter side. C) Showing there is not much variation in implict racial scores. D) Showing more variation in explitic reacial bias scores. The latter two were centred around the mean, hence, zero does not mean there is no bias."}

red_card_plot = dat %>% ggplot(aes(factor(all_reds))) +
  geom_bar() +
  xlab("Red card") +
  scale_y_continuous(name = "Count", labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey"), text = element_text(size = 10))

skin_tone_num_plot = dat %>% ggplot(aes(skin_tone_num)) +
  geom_density() +
  scale_x_continuous(name = "Skin tone rating") +
  scale_y_continuous(name = "Density", breaks = c(seq(0, 8, 2)), limits = c(0, 8)) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey"), text = element_text(size = 10))

imp_bias_plot = dat %>% ggplot(aes(imp_bias)) + 
  geom_density(adjust = 2) +
  scale_x_continuous(name = "Implicit racial bias", breaks = c(seq(-2, 2, 1)), limits = c(-2, 2)) +
  scale_y_continuous(name = "Density", breaks = c(seq(0, 40, 10)), limits = c(0, 40)) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey"), text = element_text(size = 10))

exp_bias_plot = dat %>% ggplot(aes(exp_bias)) +
  geom_density(adjust = 2) + 
  scale_x_continuous(name = "Explicit racial bias", breaks = c(seq(-2, 2, 1)), limits = c(-2, 2)) +
  scale_y_continuous(name = "Density", breaks = c(seq(0, 40, 10)), limits = c(0, 40)) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey"), text = element_text(size = 10))

plot_grid(red_card_plot, skin_tone_num_plot, imp_bias_plot, exp_bias_plot, labels = "AUTO")

@@

# Results
## Replicating *Team 23* analysis
     As mentioned in the above section the first steps were replicating the *team 23* analysis using their provided scripts. The results are reproducible without any adjustment. Skin tone has a significant effect on the odds of being send off the field ($OR = 1.311$, 95%CI [1.099, 1.563], $p = 0.003$). This means while keeping all other variables constant for every unit increase in skin tone rating (darker skin tone), the odds of being send off the field increase by roughly 131%. The interaction terms of skin tone and implicit racial bias ($OR = 0.004$, 95%CI [0.000, 23.259], $p = 0.211$) as well as skin tone and explicit bias ($OR = 1.837$, 95%CI [0.493, 6.848], $p = 0.365$) are both non-significant. In accordance with the original analysis, there are also significant differences between leagues and positions as well as implicit racial bias scores. Explicit bias scores are on verge of being non-significant. This analysis only aimed at testing reproducibility, which proved to be true, further results details are forfeited.

## Specification curve
     Shown in Figure \ref{fig:sca_plot}. The horizontal axes of the top and bottom chart show the specifications sorted based on their odds rations from lowest to highest. The odds ratios themselves are shown on the vertical axis of the top plot. The points in the top plot, which look like a line, each represent the outcome of one statistical model. Black points refers to statistical significant, red to non-significant outcomes. Overall are 89.7% of the outcomes significant. Except for the lowest outcome ($OR_{lowest} = 1.080$) are all points close to the their respective neighbours. This indicates that none of the specified models are responsible for a sudden increase in the outcome measure. The relatively smooth line suggest that sampling 1,000 specifications is, in fact, representative for all 33k models. The law of large numbers  The points can be divided into roughly three sections all non-significant ($\leq 57^{th}$), mixture of (non-)significance ($58^{th} - 247^{th}$) and all significance ($\geq 248^{th}$). The bottom table lists all covariates on its vertical axis, sorted from most positive to most negative impact. For example, player has the strongest positive impact, while club has the strongest negative impact. Each column of the table represents one mode. A coloured cell indicates the presence of the variable, an uncoloured cell its absence. Just like a above, black indicates the outcome measure was significant, red non-significance.

 The top and bottom plot work in tandem. Each outcome (i.e., point) in the top corresponds to the indicated covariates in the bottom table. Regular specification curve analyses focus on all *reasonable* specifications, hence, their number of specifications is much lower and one is, in fact, able to tell the different model specifications apart. However, the current project's goal was to define the results space, which requires all *possible* specifications. The specifications are therefore huddled extremely closely which does not allow the plot to functions in its supposed way. Thus, I create two alternative graphs. \newline

```{r sca_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Convential specification curve. Each outcome (i.e., point) in the top corresponds to the indicated covariates in the bottom table. Due to the large number of specifications, this plot is not as informative as it is supposed to be. It is, therefore, merely making an argument for needing and alternative.", warning = FALSE}

# Order by odds ratio, low to high
analysed_specifications = dat_models_specifications[order(dat_models_specifications$estimate_oddsratio), ]

# Assign identifier
analysed_specifications$id = factor(seq(1:nrow(analysed_specifications)))

# Preparing specification matrix i.e., the bottom part of the SCA plot
## Subset covariates and statistics
prepare_long_df = analysed_specifications[ , c(4:18, 20, 21, 24)]

## FALSE ro NA to better subset
prepare_long_df[prepare_long_df == FALSE] = NA

## Wide to long format
long_df = melt(setDT(prepare_long_df), id.vars = c("id", "below_alpha"), variable.name = c("estimate_oddsratio"))

## Exclude NA cases
long_df = long_df[complete.cases(long_df), ]

## Exclude outcome variable
long_df = long_df[!grepl("estimate_oddsratio", long_df$estimate_oddsratio), ]
colnames(long_df) = c("id", "below_alpha", "covariate", "x")

## Exclude value column
long_df = long_df[ , c("id", "below_alpha", "covariate")]

## order covariate effects low to high, based on the 4_covariate_effect_data.csv dataset
impact_df = dat_covariates[order(dat_covariates$estimate_oddsratio), ]

long_df$covariate = factor(long_df$covariate,
                           levels = c("club", "victories", "weight_kg", "player_cards_received", "ties", "league_country", "specific_pos", "age_yrs", "ref", "ref_cards_assigned", "goals", "height_cm", "games", "ref_country", "player"),
                           labels = c("Club", "Victories", "Weight", "Cards rec.", "Ties", "Lg. country", "Position", "Age", "Ref", "Cards assig.", "Goals", "Height", "Games", "Ref country", "Player"))

# build plot
top = analysed_specifications %>% 
  ggplot(aes(as.numeric(id), estimate_oddsratio)) +
  geom_point(aes(colour = below_alpha), size = 0.3, alpha = 0.5) +
  geom_ribbon(aes(ymin = ci_lower_oddsratio, ymax = ci_upper_oddsratio), alpha = 0.3) +
  geom_hline(yintercept = 1, color = "black") +
  scale_color_manual(values = c("red", "black")) +
  scale_x_discrete(name = "", expand = c(0.01, 0)) +
  scale_y_continuous(name = "Odds ratio", breaks = c(seq(0.9, 1.7, 0.1)), limits = c(0.89, 1.71), expand = c(0, 0)) +
  labs(title = "Specification curve", subtitle = "N = 1,000; 95% CI") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey"),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        plot.margin = unit(c(0.5, 1, 0, 1), "cm"),
        legend.position = "top",
        legend.title.align = 1,
        legend.text = element_text(size = 11),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.margin = margin(t = -0.5, b = -0.2, unit = "cm"),
        legend.box.margin = margin(t = -0.5, b = -0.2, unit = "cm"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "italic", margin = margin(0, 0, 20, 0)),
        text = element_text(size = 10)) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 1), reverse = TRUE))

bottom = long_df %>% 
  filter(id != "1") %>%
  ggplot(aes(x = id, y = covariate)) +
  geom_tile(aes(fill = below_alpha), width = 0.5, height = 0.5, color = "white") +
  scale_fill_manual(values = c("red", "black")) +
  scale_x_discrete(name = "Specifications", expand = c(0.01, 0)) +
  scale_y_discrete(name = "Covariates", expand = c(0.03, 0)) +
  theme_classic() +
  theme(legend.position = "none", 
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        plot.margin = unit(c(0, 1, 0.5, 1), "cm"),
        text = element_text(size = 10))

sca_plot = plot_grid(top, bottom, ncol = 1, align = "v")
sca_plot

## Results space
     Shown in Figure \ref{fig:rain_cloud_plot}. The horizontal axis shows the odds rations. From top to bottom the three component show the probability density distribution, the box plot and the raw data, respectively. The dashed line indicates the original results of *team 23*, which has also been the median outcome of all analytical strategies in @silberzahn2018. Based on the raw data points, the results space can be defined as the interval between 1.081 and 1.383. The 1rd quartile   \newline

```{r rain_cloud_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "The rain cloud plot is comprised of three parts. The probability density distribution, the box plot and the raw data points. This allows for an transparent and bias assessment of the results space and its distribution."}
